---
title: "assignment8"
output:
  pdf_document: default
  html_document: default
date: "2023-10-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(jsonlite)
```

making the api request and storing the results in a  dataframe
```{r}
# Define the API URL with your specific endpoint and parameters
api_url <- "https://api.nytimes.com/svc/books/v3/lists/best-sellers/history.json?api-key=8GkfuqeNdp08Sc1RtvRSxCP6DDwHgkof"

# Make a GET request to the API
response <- GET(api_url)

json_data <- content(response, "text")

json_object <- fromJSON(json_data)

best_seller <- json_object$results

head(best_seller)
```

just store the data without the na in it 
```{r}
# Assuming 'best_seller' is your original DataFrame
# Create a new DataFrame without NAs
cleaned_best_seller <- na.omit(best_seller)

```

## Data analysis 
```{r}
head(cleaned_best_seller)
```

```{r}
# Show summary statistics for numeric columns
cleaned_best_seller$price <- as.numeric(cleaned_best_seller$price)
summary(cleaned_best_seller$price)
```

```{r}
# Show unique values in categorical columns
unique_titles <- unique(cleaned_best_seller$title)
unique_publishers <- unique(cleaned_best_seller$publisher)

cat("Unique Titles:\n")
print(unique_titles)

cat("Unique Publishers:\n")
print(unique_publishers)

```

### after createting a bar plot for the distribution of publishers, it seems like Harlequin Mira has the highest distribution amongst the pther publishers, meaning that there are more books associated with that publisher
```{r}
# Create a bar plot for the 'publisher' column
barplot(table(cleaned_best_seller$publisher), main="Distribution of Publishers")

# Create a histogram for the 'price' column
hist(cleaned_best_seller$price, main="Distribution of Prices", xlab="Price")

```
### We can also note that 0 dollars is the mosty frequent price amongst all of the books being published
