---
title: "week 3 assignment"
output: html_document
date: "2023-09-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

```{r}
theUrl <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/all-ages.csv"
major <- read.table(file= theUrl, header=TRUE, sep=",")
summary(major)
```


#1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"
```{r}
filtered_major <- major %>%
  filter(grepl("DATA|STATISTICS", Major, ignore.case = TRUE))

print(filtered_major)
```

#2 Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")
```{r}
first<- c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

changed <- first

print(changed)

```

#3 Describe, in words, what these expressions will match:

(.)\1\1
"(.)(.)\\2\\1"
(..)\1
"(.).\\1.\\1"
"(.)(.)(.).*\\3\\2\\1"
**This regular expression will first match 3 identical characters consectively. The next line match any kind of word that is the same backwards as forward. The third line matches a string where the first two characters repeat at the end. The fourth line matches a 5 character string where the 1st, 3rd, and 5th characters are the same. And the 5th line matches a string that starts and ends with 3 charcters, but the last three characters is the first three characters in reverse**

#4 Construct regular expressions to match words that:

1: Start and end with the same character.
2: Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
3: Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
**1: (.)\1.*(.)\2           2:(.)\1.*(.)\2     3:(.)\1.*(.)\2.*(.)\3 **
