---
title: "Property sales analysis"
output: html_document
date: "2023-11-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RMySQL)
library(yaml)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(infer)
library(rvest)
library(mongolite)
```

Establishing a connection to the sql database
```{r}
config <- yaml::read_yaml("config.yaml")
con <- dbConnect(
  RMySQL::MySQL(),
  dbname = config$dbname,
  host = config$host,
  port = config$port,
  user = config$user,
  password =  config$password
)
```


Store each table in the database in a dataframe
```{r}
## 2018
query1 <- "SELECT * FROM 607final.2018_property_sales"
property_sales_2018 <- dbGetQuery(con, query1)

## 2019
query2 <- "SELECT * FROM 607final.2019_property_sales"
property_sales_2019 <- dbGetQuery(con, query2)

## 2020
query3 <- "SELECT * FROM 607final.2020_property_sales"
property_sales_2020 <- dbGetQuery(con, query3)

## 2022
query4 <- "SELECT * FROM 607final.2022_property_sales"
property_sales_2022 <- dbGetQuery(con, query4)
```

Store the 2021 csv dataframe in mongoddb atlas
```{r}
uri <- "mongodb+srv://bobbtilon:I8aGyfSubO4lteBf@cluster0.t25lesh.mongodb.net/cluster0?retryWrites=true&w=majority"
conn <- mongo(collection = "final", db = "cluster0", url = uri)
property_sales_2021 <-conn$find()
```

**If connection to mongodb doesn't work do the following and delete the above codechunk**
```{r}
## url <- "https://raw.githubusercontent.com/Kingtilon1/DATA607/main/Final/2021_property_sales.csv"
## property_sales_2021 <- read.csv(url)
```

## Is there a trend in the average property sales data from 2018 to 2022?
My hypothesis is that there will be a consistent increase throughout the years

### Data transformations/cleanup
```{r}
names(property_sales_2022)[names(property_sales_2022) == "﻿PropertyID"] <- "PropertyID"

clean_and_standardize <- function(df) {
  df$Sale_price <- as.numeric(gsub("[\\$,]", "", df$Sale_price))
  return(df)
}

# Assuming property_sales_2020 and property_sales_2021 are your data frames
property_sales_2020 <- clean_and_standardize(property_sales_2020)
property_sales_2021 <- clean_and_standardize(property_sales_2021)
```

### A look at the 2018 dataframe
The dataframe encompasses information on 886 properties, including categorical variables such as property type (PropType), address, and various descriptors like condominium projects (CondoProject), districts (District), neighborhoods (Nbhd), and architectural styles (Style). Numerical variables like tax keys (Taxkey) range from 1.002e+07 to 2.141e+09, with a mean value of approximately 1.394e+09. Notable summary statistics reveal a diversity of property characteristics, including the number of stories (Stories), year of construction (Year_Built), room count (Nr_of_rms), square footage (Fin_sqft), number of bedrooms (Bdrms), bathrooms (Fbath, Hbath), lot size (Lotsize), and sale price (Sale_price). Of particular interest, the sale price varies from a minimum of $2,000 to a maximum of $6,800,000, with a median value of $103,750, indicating a wide range of property values in the dataset.
```{r}
summary(property_sales_2018)
```
We see that the majority of sales occurred at around 15k which can account for things like apartments and condos to around 200,000 with a few outliers going to the million mark, and based off of the summary statistics, a sale at 6 million, but we limited the view from 0 to 1 million to see the distribution clearly.
```{r}
ggplot(property_sales_2018, aes(x = Sale_price)) +
  geom_histogram(binwidth = 10000, fill = "green", color = "black", alpha = 0.7) +
  scale_x_continuous(labels = scales::comma,  limits = c(0, 1000000)) + 
  labs(title = "Distribution of Sale Prices in 2018",
       x = "Sale Price",
       y = "Count") +
  theme_minimal()
```

### We will join all of the data sets using the bindrows function
```{r}
# Create a list of data frames
data_frames_list <- lapply(2018:2022, function(year) {
  df_name <- paste0("property_sales_", year)
  get(df_name)  
})

# Function to standardize column types
standardize_types <- function(df) {
  # Convert all columns to character
  df[] <- lapply(df, as.character)
  return(df)
}

# Combine the data frames into a single data frame
combined_df <- bind_rows(lapply(data_frames_list, standardize_types), .id = "Year")

# Calculate the average sale price for each year and convert to numeric
average_prices <- combined_df %>%
  group_by(Year) %>%
  summarise(avg_sale_price = mean(as.numeric(Sale_price), na.rm = TRUE))

combined_df <- merge(combined_df, average_prices, by = "Year")

# Create a line chart with customized x-axis labels
ggplot(average_prices, aes(x = Year, y = avg_sale_price, color = Year, group = 1)) +
  geom_line() +
  ggtitle("Average Sale Prices Over Years") +
  theme_minimal() +
  scale_x_discrete(labels = c("2018", "2019", "2020", "2021", "2022"))


```
Looking at the line plot we see that from 2018 to 2019 the average price jumped from $169,411 to $258,378 dollars, a 52.61% increase, and from 2019 to 2020 ($207,131) we see a 19.9% decrease, then from 2020 to 2021($237,485) we have a  14.66% increase and then from 2021 to 2022($271,545) we have a 14.33% increase

Lets use a linear regression model to explore the trend to confirm if there is an upward trend
```{r}
model <- lm(avg_sale_price ~ Year, data = combined_df)

# Display the summary of the regression model
summary(model)

# Visualize the regression line
ggplot(average_prices, aes(x = Year, y = avg_sale_price)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  ggtitle("Regression Line of Average Sale Prices Over Years") +
  theme_minimal()
```

## Conclusion
In this linear regression analysis, we fitted a model to explore the relationship between the year and average sale prices. The coefficients for 'Year2', 'Year3', 'Year4', and 'Year5' were found to be highly statistically significant (p-value < 0.001), indicating an upward trend in average sale prices over the years. The high R-squared values close to 1 suggest that the model effectively explains a substantial proportion of the variance in average sale prices based on the 'Year' variable. Overall, the results provide compelling evidence for a significant and positive trend in average sale prices from 2018 to 2022, implying a consistent increase in property values over this period. This confirms my hypothessi that there 
