---
title: "Assignment5"
output:
  pdf_document: default
  html_document: default
date: "2023-09-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
```


```{r}
untidy_data <- "https://raw.githubusercontent.com/Kingtilon1/DATA607/main/assignment5/flights.csv"
untidy <- data.frame(read.csv(untidy_data, header=TRUE, sep="\t"))
untidy
```

I am going to give the table a name
```{r}
colnames(untidy)[1] <- "Airline"
colnames(untidy)[2] <-"Tardiness"
```

I will now drop the row that has the NA values
```{r}
untidy <- drop_na(untidy)
```

fill the empty rows with NA
```{r}
untidy[untidy==""] <-NA
tidy <- untidy %>% fill(Airline)
tidy
```

using the pivot_longe function to reshape the data frame to a long format so that the columns from 3 to 7 are all organized under one column, as is their respective values
```{r}
tidy <- pivot_longer(tidy, cols = 3:7, names_to = "city", values_to = "Frequency")
```

replacing all the . in between the cities
```{r}
tidy$city <- str_replace_all(tidy$city, "\\.", " ")
tidy
```

By calclating the total amount of flights and the total amount of delayed, and ontime flight, I can calculate the on time rate and the delayed rate for each airline 
```{r}
airline_percentage <-
  tidy %>% 
    group_by(Airline) %>%
      summarise(
        num_flights = sum(Frequency),
        num_delays = sum(ifelse(Tardiness=='delayed', Frequency, 0))
      )

airline_percentage <- 
  airline_percentage %>%
    mutate(
        num_on_time = num_flights - num_delays,
        delay_rate = num_delays / num_flights,
        on_time_rate = (num_flights - num_delays) / num_flights
    ) %>%
      arrange(delay_rate)
airline_percentage
```


## Conclusion: AM West on time rate : 89.1%, delayed rate: 10.8%. 

## Alaska Airlines on time rate : 86.7%, delayed rate: 13.3%. 

Alaska Airlines on time rate: 86.7 percent, delayed rate 13.2%
Based off of this data, I can conclude, that on average, AM WEST Airlines is on time 3% more than Alaska airlines for a total of 89%.As well as the fact that AM WEST Airlines has a lower delay rate than Alaska Airlines which is also lower by 2 percent for a total of 11 percent if you round up. 


